SU($n$) ferromagnets near the Heisenberg critical point

### 2020-06-10 collective_sun/sun_perturbation.tex
added option to get only several low-lying states within each r-body excitation manifold added more text describing excitation manifolds

% for energy diagrams, etc.
\usepackage{tikz}
\usetikzlibrary{arrows,math}
\tikzstyle{every picture}+=[remember picture]

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta>0$ with respect to $H_0$.  Rather
than treating these excited states as ``intermediate'' or ``virtual''
in the derivation of an effective ground-state Hamiltonian, we can
perform numerical simulations in the joint Hilbert space of both
ground and excited states.  When the perturbation $\O\p{\m w,X}$ is
small, namely with operator norm
$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, performing such simulations
with an appropriate choice of excited states is equivalent to
high-order ground-state perturbation theory.  The corresponding order
of perturbation theory is determined by the choice of excited states
in a manner that will be clarified below.  Moreover, the first-class
treatment of excited states enables simulating the dynamics induced by
{\it non-perturbative} additions $\O\p{\m w,X}$ to the base
Hamiltonian $H_0$ in finite systems.

In order to simulate dynamics within excited-state manifolds, we first
need to classify the set of excited states in order to construct a
suitable basis.  In the case of SU(2), a good starting point is to
identify states with definite (non-negative) spin length
$S=\set{N/2,N/2-1,\cdots}$, and spin projection
$m_\z\in\set{S,S-1,\cdots,-S}$ onto a quantization axis.  To
generalize this classification to SU($n$), we denote the manifold of
states with definite fixed spin length $S$ by $\M_{N/2-S}$.  The PS
manifold of Dicke states with $S=N/2$ is then $\M_0$, the manifold of
single-magnon spin-wave states with $S=N/2-1$ is $\M_1$, and more
generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
Similarly, rather than the spin projection $m_\z$ we identify states
by labels $m=\p{m_\up,m_\dn}$ in which $m_\mu$ denotes the occupation
number of the single-spin state $\mu$, with $m_\up-m_\dn=m_\z$ and
$m_\up+m_\dn=N$.

The label $m$ suffices to identify the PS (Dicke) manifold
$\M_0=\oper{span}\set{\ket{m}}$, but we need additional labels to
uniquely identify states in the $r$-magnon manifold $\M_r$ with $r>0$.
Single-magnon ($r=1$) spin-wave states of SU(2) spin systems can be
written in the form $\ket{m,\m v}\equiv\O\p{\m v,\sigma_\z}\ket{m}$,
where $\m v=\p{v_1,v_2,\cdots,v_N}$ and $\sigma_\z$ is a Pauli-$z$
operator.  If an SU(2) spin system is translationally invariant, then
its spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}/\sqrt{N}$ with
wavenumber $k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2)
take the form
$\ket{m,\m v}\equiv\O\p{\m v,\sigma_\z^{\otimes r}}\ket{m}$, where now
$\m v$ is an $r$-index tensor that encodes spatial structure and
correlations in the excited state $\ket{m,\m v}$ (see Figure
\ref{fig:pair_structure}).

\begin{figure}
  \centering
  \includegraphics{pair_structure/pair_L16_A3_s1.pdf}
  \includegraphics{pair_structure/pair_L16_A3_s2.pdf}
  \includegraphics{pair_structure/pair_L16_A3_s3.pdf} \\[1em]
  \includegraphics{pair_structure/pair_L16_A3_s4.pdf}
  \includegraphics{pair_structure/pair_L16_A3_s5.pdf}
  \includegraphics{pair_structure/pair_L16_A3_s6.pdf}
  \includegraphics{pair_structure/pair_L16_A3_s7.pdf}
  \caption{Matrices $\m v_\Delta=\sum_{i,j}v^{(\Delta)}_{ij}\op{i}{j}$
    that generate two-magnon states
    $\ket{\m v_\Delta,m} \equiv \O\p{\m v_\Delta,
      \sigma_\z\otimes\sigma_\z} \ket{m}$ with definite excitation
    energy $\Delta$ in a periodic 1D chain of $N=16$ spins with
    power-law SU(2) couplings $h_{ij}=-1/\oper{dist}\p{i,j}^3$, where
    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$.  Darker
    colors correspond to larger values of $\abs*{v^{(\Delta)}_{ij}}$,
    and a blue (red) color corresponds to a positive (negative) sign
    of $v^{(\Delta)}_{ij}$.  In the case of SU($n$) with $n>2$, the
    ``two-magnon'' excitation energies $\Delta$ and matrices
    $\m v_\Delta$ remain identical, but now correspond to a degenerate
    set of states that generated by different choice of two-body
    operators in place of $\sigma_\z\otimes\sigma_\z$.}
  \label{fig:pair_structure}
\end{figure}

@@

% todo: give examples of these manifolds in SU(2)
% todo: clarify connection between multi-body excitation manifolds
%       and manifolds of fixed total spin in the SU(2) case

@@

\red{[note: text below is old and unsimplified]}

We denote the manifold of all states that can be generated by the
action of some $M$-body operator, but not any $r$-body operator with
$r<M$, by $\M_M$; we will refer to $\M_M$ as the $M$-body excitation
manifold.  We then denote the a direct sum of all $\M_r$ for $r\le M$
by $\col\M_M$, i.e.~$\col\M_M\equiv\bigoplus_{r\le M}\M_r$.  The
manifold $\col\M_M$ consists of all states accessible by the action of
$r$-body operators with $r\le M$ on PS states in $\M_0$.  These
definitions are sketched out in Figure \ref{fig:manifold_sketch}.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \draw[ultra thick] (0,0) -- node[above] {$\M_0$} ++(10em,0);
    \draw[ultra thick] (0,3em) -- node[above] {$\M_1$} ++(10em,0);
    \draw[ultra thick] (0,2*3em) -- node[above] {$\M_2$} ++(10em,0);
    \draw[->] (-.2em,0) to [bend left] (-.2em,3em);
    \node (X) at (.2em,0.5*3em) {$\hat\O_1$};
    \draw[->] (-.4em,0) to [bend left] (-.2em,2*3em);
    \node (XY) at (-2em,3em) {$\hat\O_2$};
    \draw (10em+.2em,-.2em) -- (10em+.4em,-.2em)
    -- (10em+.4em,3em+.2em) -- (10em+.2em,3em+.2em);
    \node (cM1) at (10em+1.5em,0.5*3em) {$\col\M_1$};
    \draw (10em+2.4em+.2em,-.4em) -- (10em+2.4em+.4em,-.4em)
    -- (10em+2.4em+.4em,6em+.4em) -- (10em+2.4em+.2em,6em+.4em);
    \node (cM1) at (10em+2.4em+1.5em,3em) {$\col\M_2$};
  \end{tikzpicture}
  \caption{Schematic of low-lying multi-body excitation manifolds.
    Starting with the permutationally-state manifold $\M_0$, the
    single-body excitation manifold $\M_1$ consists of states
    accessible via single-body perturbations $\hat\O_1$.  The two-body
    excitation manifold $\hat\O_2$ then consists of states accessible
    via two-body perturbations, but not single-body perturbations.
    The joint manifold $\col\M_M$ consists of all states accessible by
    operators $r$-body operators $\O_r$ with $r\le M$.}
  \label{fig:manifold_sketch}
\end{figure}

@@

Here we show that power-law SU($n$)-symmetric interactions of the form
in \eqref{eq:H_0} with $h_{ij}=-h/\abs{i-j}^\alpha$ yield a
non-vanishing many-body interaction energy gap when $\alpha\le D$,
where $D$ is the dimension of the lattice.  On a periodic lattice of
$N=L^D$ spins, the spectral gap of the interaction Hamiltonian is (see
Appendix \ref{sec:single_body_eigenstates})
\begin{align}
  \Delta
  = \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
  = h \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d \c k_{\t{SE}}}}{\abs{d}^\alpha},
\end{align}
where $\ZZ_L$ denotes the integers modulo $L$;
$k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a
singly-excited spin-wave state; and the norm $\abs{d}$ for a vector
$d$ on a periodic lattice is implicitly understood to mean the
smallest Euclidean distance of $d$ from a fixed origin.  In order to
yield the smallest excitation energy $\Delta$, the wavenumber
$k_{\t{SE}}$ should maximize the contribution of the cosine term
above, which is achieved by a wavenumber that minimizes the
oscillations of this term when integrated over the entire lattice.  A
suitable candidate for a minimal wavenumber is
$k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which corresponds to an excitation
energy
\begin{align}
  \Delta = h \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
\end{align}
Defining $\epsilon\equiv2/L$ and a rescaled domain symmetrized about
$0$, $\SS_\epsilon\simeq\ZZ_L/\epsilon$ with
$\SS_\epsilon\subset\sp{-1,1}$, we substitute $x\simeq\epsilon d$ to
get
% todo: define \simeq ?
\begin{align}
  \Delta
  = h \sum_{\substack{x\in\SS_\epsilon^D\\\abs{x}\ge\epsilon}}
  \f{1-\cos\p{\pi x_1}}{\abs{x/\epsilon}^\alpha}
  = h \epsilon^{\alpha-D} \sum_{\substack{x\in\SS_L^D\\\abs{x}\ge\epsilon}}
  \epsilon^D \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha}.
\end{align}
As $\epsilon\to0$ the discrete sum over $x$ is well approximated by an
integral that avoids an infinitesimal region about the origin, i.e.
\begin{align}
  \Delta = h \epsilon^{\alpha-D} \I_D\p{\epsilon},
  &&
  \I_D\p{\epsilon}
  \equiv \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
  \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha},
\end{align}
where the integral $\I_D\p{\epsilon}$ is defined using the interval
$\TT_a\equiv\p{-a,a}$.  The integrand of $\I_D\p{\epsilon}$ is
strictly positive and well-behaved on the entirety of its domain
except for the origin, where depending on the value of $\alpha$ the
integrand may vanish or diverge as $\abs{x}\to0$.  Together, these
facts mean that
\begin{align}
  \I_D\p{\epsilon} \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
  &&
  \Delta \stackrel{\epsilon\to0}{\sim} h \epsilon^{\alpha-D-\gamma},
\end{align}
for some $\gamma\ge0$, which implies that gap $\Delta$ is
non-vanishing when $\alpha\le D\le D+\gamma$.

@@

where

@@

is the set of all subsets (``choices'') of $q$ elements from $\ZZ_N$.

@@

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:multi_body_problem} in a basis that respects the spatial
symmetries of $H_0$.  These symmetries essentially break the set
$\D_N\p{M}$ into equivalence classes $\S_M$\footnote{For brevity, we
  will suppress the explicit dependence of $\S_M$ on $N$, as well as
  the dependence of $\S_M$ on any other factors such as lattice
  geometry.}, where elements $s\in\S_M$ are subsets
$s\subset\D_N\p{M}$ that are invariant under the appropriate symmetry
group action.  In the case of $M=2$ in a translationally invariant and
isotropic system, for example, each equivalence class $s\in\S_2$ would
consist of all pairs $\p{i,j}\in\D_N\p{2}$ that are a fixed distance
$\abs{i-j}=d_s$ apart,
i.e.~$s=\set{\p{i,j}\in\D_N\p{2}:\abs{i-j}=d_s}$.  If the tensor
$\m w$ is invariant under arbitrary permutations of its tensor
factors, this symmetry is preserved by $\hat{\m h}$, and we can place
all $k\in\D_N\p{M}$ that are equal up to permutation into the same
equivalence class.

For simplicity, we first assume that the $M$-body operator
$\O\p{\m w,X}$ obeys the same spatial symmetries as $H_0$.  For any
equivalence class $s\in\S_M$, we then define a uniform superposition
$\ket{s}$ over all $k\in s$, as well as a projector $\P_\S$ onto the
space of uniform superpositions:
\begin{align}
  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
  &&
  \P_{\S_M} \equiv \sum_{s\in\S_M} \op{s},
\end{align}
where $\abs{s}$ is the number of elements in $s$.  In turn, we expand
a vector of $M$-body operator coefficients
$\m w=\sum_{k\in\D_N\p{M}}w_k\ket{k}$ in this symmetrized basis:
\begin{align}
  \m w = \P_{\S_M} \m w
  = \sum_{s\in\S_M} \sum_{k\in s} w_k \ket{s} \bk{s|k}
  = \sum_{s\in\S_M} w_s \sqrt{\abs{s}} \ket{s},
\end{align}
where all $w_k$ with $k\in s$ are equal by assumption, so we can
define $w_s\equiv w_k$ for any $k\in s$.  Finally, we expand the
projection of $\hat{\m h}$ onto the symmetric subspace:
\begin{align}
  \P_{\S_M} \hat{\m h} \P_{\S_M}
  = \sum_{s\in\S_M} \f{\ket{s}}{\sqrt{\abs{s}}} \sp{
    \sum_{\substack{k\in s\\a\in\ZZ_M}}
    \sum_{\substack{p\in\ZZ_N\\p\notin k}}
    h_{p k_a} \f{\bra{\sp{k_{a:p}}}}{\sqrt{\abs{\sp{k_{a:p}}}}}
    + \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b}
    \f{\bra{\sp{k_{a\lra b}}}}{\sqrt{\abs{\sp{k_{a\lra b}}}}}
    - h^{\t{D}}_s \f{\bra{s}}{\sqrt{\abs{s}}}},
\end{align}
where $h^{\t{D}}_s\equiv h^{\t{D}}_k$ for any $k\in s$, with
$h^{\t{D}}_k$ defined in \eqref{eq:multi_body_op_diag_element}; and
$\sp{k}$ denotes the equivalence class of $k$.  The reduced multi-body
eigenvalue problem is then
\begin{align}
  \P_{\S_M} \hat{\m h} \P_{\S_M} \c \m w = \Delta \m w.
\end{align}
More generally, even if the couplings $\m w$ of $\O\p{\m w,X}$ do not
obey the spatial symmetries as $H_0$, these symmetries still endow
$\hat{\m h}$ with a block-diagonal structure that reduces the
complexity of the multi-body eigenvalue problem.  We leave this
reduction to future work.  % todo: perform reduction

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The general case: a diagrammatic expansion}

>>>>

% for energy diagrams, etc.
\usepackage{tikz}
\usetikzlibrary{arrows,math,calc}
\tikzstyle{every picture}+=[remember picture]
\tikzset{
    thickest/.style={line width=3pt}
}

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta>0$ with respect to $H_0$.  Rather
than treating these excited states as ``intermediate'' or ``virtual''
in the derivation of an effective ground-state Hamiltonian, we can
perform numerical simulations in the joint Hilbert space of both
ground and excited states.  When the perturbation $\O\p{\m w,X}$ is
small, namely with operator norm
$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are
equivalent (up to perturbative corrections) to a higher-order
perturbation theory.  Moreover, the first-class treatment of excited
states in finite systems enables simulating the dynamics induced by
{\it non-perturbative} additions $\O\p{\m w,X}$ to the base
Hamiltonian $H_0$.

In order to simulate dynamics within excited-state manifolds, we first
need to classify the set of excited states in order to construct a
suitable basis.  In the case of SU(2), a common starting point is to
identify states with a definite (non-negative) spin length
$S=\set{N/2,N/2-1,\cdots}$, and a spin projection
$m_\z\in\set{S,S-1,\cdots,-S}$ onto a quantization axis.  To
generalize this classification to SU($n$), we denote the manifold of
states with definite fixed spin length $S$ by $\M_{N/2-S}$.  The PS
manifold of Dicke states with $S=N/2$ is then $\M_0$, the manifold of
single-magnon spin-wave states with $S=N/2-1$ is $\M_1$, and more
generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
Rather than the spin projection $m_\z$ we then identify states by
labels $m=\p{m_\up,m_\dn}\in\A_2\p{N}$ in which $m_\mu$ denotes the
occupation number of the single-spin state $\mu$, with
$m_\up-m_\dn=m_\z$ and $m_\up+m_\dn=N$.

The label $m$ suffices to identify the PS (Dicke) manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_2\p{N}}$, but we need additional
labels to uniquely identify states in the $r$-magnon manifold $\M_r$
with $r>0$.  Single-magnon ($r=1$) spin-wave states of SU(2) spin
systems can be written in the form
\begin{align}
  \ket{m,\m v} \equiv \O\p{\m v,\sigma_\z} \ket{m}
  = \sum_j v_j \sigma_\z^{(j)} \ket{m},
\end{align}
where each $v_j$ is a scalar and $\sigma_\z$ is a Pauli-$z$ operator.
If an SU(2) spin system is translationally invariant, then these
spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}/\sqrt{N}$ with
wavenumber $k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2)
take the form
$\ket{m,\m v}\equiv\O\p{\m v,\sigma_\z^{\otimes r}}\ket{m}$, where now
$\m v$ is an $r$-index tensor that encodes the structure of
correlations in the excited state $\ket{m,\m v}$ (see Figure
\ref{fig:structure}).  Note that $M$-body operators will generally
couple the PS manifold $\M_0$ to all $r$-magnon manifolds $\M_r$ with
$r\le M$ (see Figure \ref{fig:manifold_sketch}).  To generalize the
notion of $r$-magnon manifolds $\M_r$ to the case of SU($n$), where
there is no obvious notion of a spin length $S$, we therefore define
$\M_r$ for $r>0$ as the manifold of states that (i) can be generated
by the action of $r$-body operators on $\M_0$, and (ii) are orthogonal
to all states in $\bigoplus_{s<r}\M_s$.

\begin{figure}
  \centering
  \includegraphics{structure/mat_L101_a2.pdf}
  \caption{Visualizations of dimension-$r$ tensors $\m v$ that
    generate the lowest-lying $r$-magnon states
    $\ket{\m v,m} \equiv \O\p{\m v, \sigma_\z^{\otimes{r}}} \ket{m}$
    with definite excitation energy $\Delta$ above the PS manifold
    $\M_0$ of a periodic 1D lattice of $N=101$ spins interacting via
    the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
    couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
    minimal distance between spins $i,j$ on a periodic 1D lattice.
    Top row ($r=2$) shows the full 2D matrix
    $\m v = \sum_{i,j}v_{ij}\op{i}{j}$, and bottom row ($r=3$) shows a
    single 2D slice of the full 3D tensor,
    $\tilde{\m v}_0 = \sum_{i,j}v_{ij,0}\op{i}{j}$.  Blue (red) pixels
    correspond to matrix elements with a positive (negative) sign, and
    darker colors indicate larger absolute values.  In the case of
    $r=2$, translational invariance enforces $v_{ij}=v_{i-d,j-d}$ for
    all displacements $d$, so every row of $\m v$ is some translation
    of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
    of the full 3D tensor can be obtained by appropriately translating
    the 2D slice $\tilde{\m v}_0$ shown above.  In the case of SU($n$)
    with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
    tensors $\m v$ remain identical, but now correspond to a
    degenerate set of states that are generated by different choices
    of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
  \label{fig:structure}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \draw[thickest] (0,0) -- ++(10em,0);
    \foreach \hh in {0.15em,0.05em,-0.05em,-0.15em} {
      \draw (0,3em+\hh) -- ++(10em,0);
    }
    \draw[->] (7em,.25em) arc (220:0:.4em);
    \draw[->] (7em,.25em) to [bend left]
    node[left] {$\O\p{\m v,\sigma_\z}$} (7em,2.7em);
    \node () at (5em,4.7em) {$\vdots$};
    \foreach \hh in {0.25em,0.15em,0.05em,-0.05em,-0.15em,-0.25em} {
      \draw (0,6em+\hh) -- ++(10em,0);
    }
    \draw[->] (-.4em,0em) arc (220:0:.4em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,3em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,4.3em);
    \draw[->] (-.4em,0) to [bend left, in=130]
    node[left] {$\O\p{\m w,\sigma_\z^{\otimes M}}$} (-.2em,6em);
    \node[right] () at (10em,0) {$\M_0,S=\frac{N}{2}$};
    \node[right] () at (10em,3em) {$\M_1,S=\frac{N}{2}-1$};
    \node[right] () at (10em,6em) {$\M_M,S=\frac{N}{2}-M$};
  \end{tikzpicture}
  \caption{Schematic of relationships between $r$-magnon manifolds
    $\M_r$ with total spin $S=N/2-r$ in an SU(2) spin system.
    Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the
    ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$
    as well as $\M_0$ itself, while $M$-body operators
    $\O\p{\m w,\sigma_\z^{\otimes M}}$ couple $\M_0$ to all $r$-magnon
    manifolds $\M_r$ with $r\le M$.}
  \label{fig:manifold_sketch}
\end{figure}

@@

@@

\red{[note: the remainder of this section is old, and will be replaced
  entirely]}

@@

Here we show that in $D$ spatial dimensions, power-law
SU($n$)-symmetric interactions of the form in \eqref{eq:H_0} with
$h_{ij}=-1/\abs{i-j}^\alpha$ yield a non-vanishing spectral gap when
$\alpha\le D$.  For definiteness, we will consider a periodic
hypercubic lattice of $N=L^D$ spins.  The spectral gap of the
interaction Hamiltonian is then (see Appendix
\ref{sec:single_body_eigenstates})
\begin{align}
  \Delta
  = \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
  = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d \c k_{\t{SE}}}}{\abs{d}^\alpha},
\end{align}
where $\ZZ_L$ denotes the set of integers modulo $L$;
$k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a single-magnon
spin-wave state; and the norm $\abs{d}$ for a vector $d$ on a periodic
lattice is implicitly understood to mean the smallest Euclidean
distance of $d$ from a fixed origin.  In order to yield the smallest
excitation energy $\Delta$, the wavenumber $k_{\t{SE}}$ should
maximize the contribution of the cosine term above, which is achieved
by a wavenumber that minimizes the oscillations of this term when
integrated over the entire lattice.  A suitable candidate for a
minimal wavenumber is $k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which
corresponds to an excitation energy
\begin{align}
  \Delta = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
\end{align}
Defining $\epsilon\equiv2/L$ and a rescaled domain symmetrized about
$0$, $\SS_\epsilon\simeq\ZZ_L/\epsilon$ with
$\SS_\epsilon\subset\sp{-1,1}$, we substitute $x\simeq\epsilon d$ to
get
\begin{align}
  \Delta
  = \sum_{\substack{x\in\SS_\epsilon^D\\\abs{x}\ge\epsilon}}
  \f{1-\cos\p{\pi x_1}}{\abs{x/\epsilon}^\alpha}
  = \epsilon^{\alpha-D} \sum_{\substack{x\in\SS_L^D\\\abs{x}\ge\epsilon}}
  \epsilon^D \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha}.
\end{align}
As $\epsilon\to0$ the discrete sum over $x$ is well approximated by an
integral that avoids an infinitesimal region about the origin, i.e.
\begin{align}
  \Delta = \epsilon^{\alpha-D} \I_D\p{\epsilon},
  &&
  \I_D\p{\epsilon}
  \equiv \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
  \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha},
\end{align}
where the integral $\I_D\p{\epsilon}$ is defined using the symmetric
interval $\TT_a\equiv\p{-a/2,a/2}$.  The integrand of
$\I_D\p{\epsilon}$ is strictly positive and well-behaved on the
entirety of its domain except for the origin, where depending on the
value of $\alpha$ the integrand may vanish or diverge as
$\abs{x}\to0$.  Together, these facts mean that
\begin{align}
  \I_D\p{\epsilon} \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
  &&
  \Delta \stackrel{\epsilon\to0}{\sim} \epsilon^{\alpha-D-\gamma},
\end{align}
for some $\gamma\ge0$, which implies that gap $\Delta$ is
non-vanishing when $\alpha\le D\le D+\gamma$.

@@

where for convenience we define

@@

as the set of all subsets (``choices'') of $q$ elements from $\ZZ_N$.

@@

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:multi_body_problem} in a basis that respects the spatial
symmetries of $H_0$.  These symmetries essentially break the set
$\D_N\p{M}$ into equivalence classes $\S_M$\footnote{For brevity, we
  will suppress the explicit dependence of $\S_M$ on $N$, as well as
  other factors such as lattice geometry.}, where elements $s\in\S_M$
are subsets $s\subset\D_N\p{M}$ that are invariant under the
appropriate symmetry group action.  In the case of $M=2$ in a
translationally invariant and isotropic system, for example, each
equivalence class $s\in\S_2$ would consist of all pairs
$\p{i,j}\in\D_N\p{2}$ that are a fixed distance $\abs{i-j}=d_s$ apart,
i.e.~$s=\set{\p{i,j}\in\D_N\p{2}:\abs{i-j}=d_s}$.  If the tensor
$\m w$ is invariant under arbitrary permutations of its tensor
factors, this symmetry is preserved by $\hat{\m h}$, and we can place
all $k\in\D_N\p{M}$ that are equal up to permutation into the same
equivalence class.

For simplicity, we first assume that the $M$-body operator
$\O\p{\m w,X}$ obeys the same spatial symmetries as $H_0$.  For any
equivalence class $s\in\S_M$, we then define a uniform superposition
$\ket{s}$ over all $k\in s$, as well as a projector $\P_\S$ onto the
space of uniform superpositions:
\begin{align}
  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
  &&
  \P_{\S_M} \equiv \sum_{s\in\S_M} \op{s},
\end{align}
where $\abs{s}$ is the number of elements in $s$.  In turn, we expand
a vector of $M$-body operator coefficients
$\m w=\sum_{k\in\D_N\p{M}}w_k\ket{k}$ in this symmetrized basis:
\begin{align}
  \m w = \P_{\S_M} \m w
  = \sum_{s\in\S_M} \sum_{k\in s} w_k \ket{s} \bk{s|k}
  = \sum_{s\in\S_M} w_s \sqrt{\abs{s}} \ket{s},
\end{align}
where all $w_k$ with $k\in s$ are equal by assumption, so
$w_s\equiv\sum_{k\in s}w_k/\abs{s}=w_\ell$ for any $\ell\in s$.
Finally, we expand the projection of $\hat{\m h}$ onto the symmetric
subspace:
\begin{align}
  \P_{\S_M} \hat{\m h} \P_{\S_M}
  = \sum_{s\in\S_M} \f{\ket{s}}{\sqrt{\abs{s}}} \sp{
    \sum_{\substack{k\in s\\a\in\ZZ_M}}
    \sum_{\substack{p\in\ZZ_N\\p\notin k}}
    h_{p k_a} \f{\bra{\sp{k_{a:p}}}}{\sqrt{\abs{\sp{k_{a:p}}}}}
    + \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b}
    \f{\bra{\sp{k_{a\lra b}}}}{\sqrt{\abs{\sp{k_{a\lra b}}}}}
    - h^{\t{D}}_s \f{\bra{s}}{\sqrt{\abs{s}}}},
\end{align}
where $h^{\t{D}}_s\equiv h^{\t{D}}_k$ for any $k\in s$, with
$h^{\t{D}}_k$ defined in \eqref{eq:multi_body_op_diag_element}; and
$\sp{k}$ denotes the equivalence class of $k$.  The reduced multi-body
eigenvalue problem is then
\begin{align}
  \P_{\S_M} \hat{\m h} \P_{\S_M} \c \m w = \Delta \m w.
\end{align}
More generally, even if the couplings $\m w$ of $\O\p{\m w,X}$ do not
obey the spatial symmetries as $H_0$, these symmetries still endow
$\hat{\m h}$ with a block-diagonal structure that reduces the
complexity of the multi-body eigenvalue problem.  We leave this
reduction to future work.  % todo: perform reduction

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A diagrammatic expansion}
